Should geophysicists use gravity anomaly or gravity disturbance?

### 2018-05-18 manuscript/letter.tex
Review the conclusions
Removed a lot of repetition and added a paragraph to the end to end in a
high note with a recommendation of what people can do (just stop using
orthometric heights).

Although the gravity disturbance is a well-known quantity in geodesy,
its use in the applied geophysics is uncommon.
For years, geophysicists have been using gravity anomaly for modelling
the gravitational effects due to geological bodies.
The question "Should geophysicists use gravity anomaly or gravity
disturbance for modelling the gravitational effect of gravity sources?"
is of more than academic interest.

Theoretically, the gravity anomaly is the difference between the
gravity on the Geoid and the normal gravity on the reference ellipsoid,
both at the same geodetic latitude and longitude.
On the other hand, the gravity disturbance is defined by the difference
between gravity and normal gravity at the same point.
Because these quantities are at the same point, the gravity disturbance
is a purely gravitational effect due to the gravity sources whereas the
gravity anomaly is a combination of gravitational and centrifugal
effects.
Consequently, the gravity disturbance is a harmonic function whereas the
gravity anomaly is not.
Because the gravity disturbance is a purely gravitational effect, it can
be represented by a harmonic function and satisfies the premise required
by the most techniques for processing potential-field data (e.g.,
upward/downward continuation, data processing with equivalent layer,
conversions between gravity and magnetic data, computation of vertical
derivatives via Fourier and Hilbert transforms).
Hence, the gravity disturbance is the more appropriated quantity for
representing the gravity effect produced by the gravity sources.

In practical problems of applied geophysics, geophysicists perform the
gravity-forward modelling by computing the vertical component of the
gravitational attraction exerted by the sources at the observation
points, regardless of the Geoid or reference ellipsoid positions,
for example.
This means that, actually, they compute gravity disturbance to fit
a gravity anomaly (e.g., Bouguer anomaly) and this is a clear inconsistency.
Moreover, upward or downward continuations, for example, are commonly applied to
gravity anomalies. In this case, the gravity anomaly is wrongly considered a
harmonic function and this is another inconsistency.

By adopting these practical procedures, geophysicists are implicitly
taking the gravity anomaly as a good approximation of the gravity
disturbance. But why should geophysicists care about this usual
procedure? Because it may result in wrong interpretations of
quantitative nature. In Brazil, for example, the maximum difference
between the gravity disturbance and the free-air anomaly reaches 9.3
mGal. Such difference on the Earth's surface may lead, for example, to
an error of about xxx m in the depth-to-center estimate of a sphere-like
source with true depth-to-center at xxx m, radius xxxx m and density
contrast of xxx $kg \, m^{-3}$.

>>>>

Although the gravity disturbance is well-known in geodesy,
its use in applied geophysics is uncommon.
For years, geophysicists have been using gravity anomalies for modelling
the gravitational effects of geological bodies.
Wether geophysicists should use the gravity anomaly or the gravity disturbance
is of more than mere academic interest.
It is a question that underpins the very assumptions about what it is that we
are actually modelling.

The gravity anomaly is the difference between
gravity on the geoid and normal gravity on the reference ellipsoid.
The gravity disturbance is the difference between gravity and normal gravity at
the same point.
Because of this, the gravity disturbance is a purely gravitational effect,
whereas the gravity anomaly is a combination of gravitational and centrifugal
effects.
Consequently, the gravity disturbance is a harmonic function while the
gravity anomaly is not.
Being a harmonic function is the premise required by most techniques for
processing potential-field data
(e.g., upward/downward continuation, operations with the equivalent layer,
conversions between gravity and magnetic data, computation of vertical
derivatives).
Hence, the gravity disturbance is the logical choice for
representing the gravitational effects produced by the heterogeneous density
distribution of the Earth.

More often than not, gravity modelling is done by computing the
vertical component of the gravitational acceleration exerted by the sources.
We have shown that, in doing so, one is actually computing the gravity
disturbance while attempting to fit a gravity anomaly.
This is a clear inconsistency.
Doing so may even result in wrong interpretations of a quantitative nature.
In Brazil, for example, the maximum difference
between the gravity disturbance and the free-air anomaly reaches 9 mGal.
Such difference at the Earth's surface may lead to
an error of XXX m in the depth-to-center estimate of a sphere with
true depth-to-center at xxx m, radius xxxx m, and density
contrast of xxx $kg.m^{-3}$.

Fortunately, all of these issues are nullified during data processing by
substituting orthometric heights (i.e., with respect to the geoid) with
geometric heights (i.e., with respect to the ellipsoid).
For modern acquisitions, this requires no extra work because GPS already
provide geometric heights.
The free-air correction can even be bypassed entirely.
As a bonus, the gravity disturbance is conceptually simpler to understand and
doesn't require the concept of the geoid with which learners often struggle.
